Selecting Computations: Theory and Applications

### 2012-04-03 paper/vago.uai/blinkered.tex
as submitted

The myopic policy is an extreme approximation, often stopping significantly early.
For a natural class of metalevel probability models we can define a better approximation:

\begin{dfn}\label{dfn:independent-actions}
	A metalevel probability model $\mathcal{M}=(U_1,\dots,U_k,\Evidence)$ 
	has \term{independent actions} if the computational variables can be partitioned 
	$\Evidence = \Evidence_1\cup\dots\cup\Evidence_k$ such that such that
	the sets $\{U_i\}\cup\Evidence_i$ are independent of each other for different $i$.	
\end{dfn}

@@

This is a better approximation than the myopic policy as it takes into account more possible future strategies.
Indeed, $Q^m(s,a) \le Q^b(s,a) \le Q^*(s,a)$.  Intuitively, whereas the myopic policy can only see one 
computation ahead, the blinkered policy can see arbitrary far ahead but with blinkers on so can only see
the computations involving a single action at a time.

\begin{dfn}\label{dfn:one-action}
	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$ with independent actions,
	a \term{one-action metalevel decision problem} for $i=1,\dots,k$ is the metalevel decision
	problem $M^1_{i,m} = (S_i,s_0,A_{s0},T_i,R_i)$ defined by the metalevel probability
	model $(U_0,U_i,\Evidence_i)$ with $U_0=m$.
\end{dfn}

Note that given that state $s$ of a metalevel decision problem, we can form a state
$s_i$ by taking only the results of computations in $\Evidence_i$ (see \dfnref{dfn:metalevel-mdp}).
By action independence, $\mu_i(s)$ is a function only of $s_i$.

\begin{thm}\label{thm:blinkered}
Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$ with independent actions,
let $M^1_{i,\mu}$ be the $i$th one-action metalevel decision problem for $i=1,\dots,k$.
Then for any $s\in S$, whenever $E_i\in A_s\cap\Evidence_i$ we have:
\[
	Q^b_M(s,E_i) = Q^*_{M^1_{i,m_i}}(s_i, E_i)
\]
where $m_i = \max_{j\neq i} \mu_i(s)$.
\end{thm}

\begin{comment}
	\begin{proof}
	Fix a state $s$, a $E_i\in A_s$ and take any $\pi\in\Pi^b_i$.  Note that such policies
	are equivalent to polices $\pi'$ on $M^1_{1,m}$, and all such policies are represented.
	Consider $Q^\pi(s,E_i)$.  As $\pi(s)\in\Evidence_i$ for all $s\in S$, by action independence $\mu_j(S_n) = \mu_j(s)$.
	By this and \thmref{thm:value-of-computation}, then,
	\[
		Q^\pi(s,E_i) = \IE^\pi_M[ -c\,N + \max(\mu_i(S_N), m_i) \given S_0=s, A_0=E_i].
	\]
	Noting that $\mu_i(S_N)$ is a function only of $(S_N)_i$, and that since 
	But then this is exactly $Q^{\pi'}_{M^1_{i,m_i}}(s_i, E_i)$.  Taking the supremum
	over $\pi$ gives the result.
	\end{proof}	
\end{comment}

\thmref{thm:blinkered} shows that to compute the blinkered policy we need
only compute the optimal policies for all $i$th one-action metalevel decision problems.

For the Bernoulli problem with $k$ actions, the one-action metalevel decision problems
are all one-action Bernoulli problems (\exampleref{example:bernoulli}).  By \thmref{thm:one-action-bound}
these policies perform at most $1/4c - 3$ computations.
As a result, the blinkered policy can be numerically computed in time $O(D/c^2)$ 
independent of $k$ by backwards induction, where $D$ is the number of points $m\in[0,1]$
we compute $Q^*_{M^1_{i,m}}(s)$ for\footnote{in our experiments below, $D=129$ points are equally spaced,
using linear interpolation between points}.  This will be worth the cost in 
the same situations as mentioned at the of \secref{sec:optimal}.

\figref{fig:blinkered} compared the expected regret of the blinked policy on the 
Bernoulli sampling problem with $k=25$ and varying cost of samples to several 
alternatives, where the regret include the cost of sampling:
\[
	R = (\max_i U_i) - U_j + c\,n
\]
if the policy performs $n$ computations and selects action $j$.
Myopic is the myopic policy for Bernoulli sampling. ESPb is the 
policy recently proposed by \citet{Chick+Frazier:2011}, which approximates the 
blinkered policy for sampling normal distributions by moving to continuous time
and numerically solving a diffusion equation.  UCB1-B uses UCB$(1/\sqrt2)$ \cite{Auer+et+al:2002} to
choose which actions to sample and the blinkered policy to stop.  UCB1-b uses
UCB$(1/\sqrt2)$ to sample, is given the average number of samples blinkered
uses for the given cost.

The blinkered policy significantly outperforms all others.  The myopic policy
plateaus as it quickly reaches a position where no single computation can
change the final decision between actions.  ESPb performs quite well given
that is making a normal approximation to the Beta posterior.  
UCB1-B and UCB1-b's curve shows that even given a good stopping rule, UCB1's
choice of actions to sample can be significantly improved on.

\begin{figure}[htb]
\centering
\includegraphics[scale=0.7, trim=90 70 400 300]{blinkered-regret.pdf}
\caption{Average regret of various policies as a function of the cost in 
a 25-action Bernoulli sampling problem.  Averaged over 1000 trials.}
\label{fig:blinkered}
\end{figure}

\begin{comment}
set xlabel "Cost"
set ylabel "Regret"

set terminal postscript enhanced linewidth 2
set output "blinkered-regret.ps"
set size 0.5, 0.5

set logscale xy

set key left bottom
plot [0.1:0.0005] [0.05:1] "blinkered-regret.dat" title "Blinkered" with lines lw 2, "myopic-regret.dat" title "Myopic" with lines lw 2, "espb-regret.dat" title "ESPb" with lines lw 2, "ucb-blinkered-regret.dat" title "UCB1-B" with lines lw 2, "ucb-blinkered-table-regret.dat" with lines lw 2 title "UCB1-b"
\end{comment}

>>>>

The myopic policy is an extreme approximation, often stopping far too early.
A better approximation can be obtained, at least for the case where
each computation can only affect the value of one action.
The technical definition (closely related to {\em subtree independence} in Russell and Wefald's work) is
as follows:

\begin{dfn}\label{dfn:independent-actions}
	A metalevel probability model $(U_1,\dots,U_k,\Evidence)$ 
	has \term{independent actions} if the computational variables can be partitioned 
	$\Evidence = \Evidence_1\cup\dots\cup\Evidence_k$ such that such that
	the sets $\{U_i\}\cup\Evidence_i$ are independent of each other for different $i$.	
\end{dfn}

With independent actions, we can talk about metalevel policies that focus on
computations affecting a single action. These policies are not myopic---they can consider arbitrarily many computations---but they are {\em blinkered} because they can look
in only a single direction at a time:

@@

Clearly, blinkered policies are better than myopic: $Q^m(s,a) \le Q^b(s,a) \le Q^*(s,a)$.
Moreover, the blinkered policy can be computed in time proportional to the number of arms, by breaking the
decision problem into separate subproblems:

\begin{dfn}\label{dfn:one-action}
	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$ with independent actions,
	a \term{one-action metalevel decision problem} for $i=1,\dots,k$ is the metalevel decision
	problem $M^1_{i,\lambda} = (S_i,s_0,A_{s0},T_i,R_i)$ defined by the metalevel probability
	model $(U_0,U_i,\Evidence_i)$ with $U_0=\lambda$.
\end{dfn}

Note that given a state $s$ of a metalevel decision problem, we can form a state
$s_i$ by taking only the results of computations in $\Evidence_i$ (see \dfnref{dfn:metalevel-mdp}).
By action independence, $\mu_i(s)$ is a function only of $s_i$.

\begin{thm}\label{thm:blinkered}
Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$ with independent actions,
let $M^1_{i,\lambda_i}$ be the $i$th one-action metalevel decision problem for $i=1,\dots,k$.
Then for any $s\in S$, whenever $E_i\in A_s\cap\Evidence_i$ we have:
\[
	Q^b_M(s,E_i) = Q^*_{M^1_{i,\mu^*_{-i}}}(s_i, E_i)
\]
where $\mu^*_{-i} = \max_{j\neq i} \mu_j(s)$.
\end{thm}

\begin{hiddenproof}
	\begin{proof}
	Fix a state $s$, a $E_i\in A_s$ and take any $\pi\in\Pi^b_i$.  Note that such policies
	are equivalent to polices $\pi'$ on $M^1_{1,m}$, and all such policies are represented.
	Consider $Q^\pi(s,E_i)$.  As $\pi(s)\in\Evidence_i$ for all $s\in S$, by action independence $\mu_j(S_n) = \mu_j(s)$.
	By this and \thmref{thm:value-of-computation}, then,
	\[
		Q^\pi_M(s,E_i) = \IE^\pi_M[ -c\,N + \max(\mu_i(S_N), m_i) \given S_0=s, A_0=E_i].
	\]
	Noting that $\mu_i(S_N)$ is a function only of $(S_N)_i$, and that since 
	But then this is exactly $Q^*_{M^1_{i,\mu^*_{-i}}}(s_i, E_i)$.  Taking the supremum
	over $\pi$ gives the result.
	\end{proof}	
\end{hiddenproof}

\thmref{thm:blinkered} shows that to compute the blinkered policy we need
only compute the optimal policies for $k$ separate one-action problems.

For the Bernoulli problem with $k$ actions, the one-action metalevel decision problems
are all one-action Bernoulli problems (\exampleref{example:bernoulli}).  By \thmref{thm:one-action-bound}
these policies perform at most $1/4c - 3$ computations.
As a result, the blinkered policy can be numerically computed in time $O(D/c^2)$ 
independent of $k$ by backwards induction, where $D$ is the number of points $\lambda\in[0,1]$
for which we compute $Q^*_{M^1_{i,m}}(s)$.\footnote{in our experiments below, $D=129$ points are equally spaced,
using linear interpolation between points.}  This will be worth the cost in 
the same situations as mentioned at the end of \secref{sec:optimal}.

\figref{fig:blinkered} compares the blinkered policy 
to several other policies from the literature, using a
Bernoulli sampling problem with $k=25$ and a wide range of values for the step cost $c$.
Performance is measured by expected {\em regret}, where the regret includes the cost of sampling:
$R = (\max_i U_i) - U_j + c\,n$
where $n$ is the number of computations and $j$ is the action actually selected.
The blinkered policy significantly outperforms all others.  The myopic policy
plateaus as it quickly reaches a position where no single computation can
change the final action choice.  ESPb performs quite well given
that is making a normal approximation to the Beta posterior.  
The curves for UCB1-B and UCB1-b show that even given a good stopping rule, UCB1's
choice of actions to sample is not ideal.

\begin{figure}[htb]
\centering
\includegraphics[scale=0.7, trim=90 70 400 300]{blinkered-regret.pdf}
\caption{Average regret of various policies as a function of the cost
in a 25-action Bernoulli sampling problem, over 1000 trials. Error bars
omitted as they are negligible (the relative error is at most 0.03).}
\label{fig:blinkered}
\end{figure}

\begin{comment}
# N: In ~/research/src/mlrl

set terminal postscript enhanced linewidth 2

set size 0.5, 0.5
set logscale xy

set xlabel "Cost"
set ylabel "Samples"
set key left top

set output "blinkered-steps.ps"

plot [0.1:0.00047] "blinkered-steps.dat" title "Blinkered" with lines lw 2, "myopic-steps.dat" title "Myopic" with lines lw 2, "espb-steps.dat" title "ESPb" with lines lw 2, "ucb-blinkered-steps.dat" title "UCB1-B" with lines lw 2, "ucb-blinkered-table-steps.dat" with lines lw 2 title "UCB1-b"

set xlabel "Cost"
set ylabel "Regret"
set key left bottom

set output "blinkered-regret.ps"

plot [0.1:0.00047] [0.05:1] "blinkered-regret.dat" title "Blinkered" with lines lw 2, "myopic-regret.dat" title "Myopic" with lines lw 2, "espb-regret.dat" title "ESPb" with lines lw 2, "ucb-blinkered-regret.dat" title "UCB1-B" with lines lw 2, "ucb-blinkered-table-regret.dat" with lines lw 2 title "UCB1-b"

set output "blinkered-regret-linespoints.ps"

plot [0.1:0.00047] [0.05:1] "blinkered-regret.dat" title "Blinkered" with linespoints lw 2, "myopic-regret.dat" title "Myopic" with linespoints lw 2, "espb-regret.dat" title "ESPb" with linespoints lw 2, "ucb-blinkered-regret.dat" title "UCB1-B" with linespoints lw 2, "ucb-blinkered-table-regret.dat" with linespoints lw 2 title "UCB1-b"

set output "blinkered-regret-errorlines.ps"

plot [0.1:0.00047] [0.05:1] "blinkered-regret2.dat" title "Blinkered" with errorlines lw 2, "myopic-regret2.dat" title "Myopic" with errorlines lw 2, "espb-regret2.dat" title "ESPb" with errorlines lw 2, "ucb-blinkered-regret2.dat" title "UCB1-B" with errorlines lw 2, "ucb-blinkered-table-regret2.dat" with errorlines lw 2 title "UCB1-b"

set output "blinkered-regret-errorlines-nopoints.ps"

plot [0.1:0.00047] [0.05:1] "blinkered-regret.dat" notitle with lines lw 2, "myopic-regret.dat" notitle with lines lw 2, "espb-regret.dat" notitle with lines lw 2, "ucb-blinkered-regret.dat" notitle with lines lw 2, "ucb-blinkered-table-regret.dat" with lines lw 2 notitle, "blinkered-regret2.dat" title "Blinkered" with errorbars lw 2, "myopic-regret2.dat" title "Myopic" with errorbars lw 2, "espb-regret2.dat" title "ESPb" with errorbars lw 2, "ucb-blinkered-regret2.dat" title "UCB1-B" with errorbars lw 2, "ucb-blinkered-table-regret2.dat" with errorlines lw 2 title "UCB1-b"
\end{comment}

### 2012-04-03 paper/vago.uai/context.tex
as submitted

\begin{example}[Non-indexability]
Consider a metalevel probability model with three actions. 
	$U_0=m$ has a known value (the context),
	$U_1$ is equally likely to be -1.5 or 1.5 (low mean, high variance),
	and $U_2$ is equally likely to be 0.25 or 1.75 (high mean, low variance).
The two computations are to observe exactly $U_1$ and $U_2$, respectively, with a cost 0.2 of computation.
The corresponding metalevel MDP has 9 states and can be solved exactly.
\figref{fig:swap-counterex} plots $Q^*_m(s_0,U_i) - Q^*_m(s_0,\bot)$ as a function of the known value $m$.
As the context $m$ varies the optimal action inverts from observing 1 to observing 2.
Inversions like this are impossible for index policies.
\end{example}

\begin{figure}[htb]
\centering
\includegraphics[scale=0.7, trim=90 70 400 300]{swap-counterex.pdf}
\caption{Optimal Q-values as a function of the utility of the fixed alternative.  Note the inversion
where for low $m$ observing action~$1$ is strictly optimal, while for medium $u$ observing action~$2$ is strictly optimal.}
\label{fig:swap-counterex}
\end{figure}

@@

set terminal postscript enhanced linewidth 2
set size 0.5,0.5
set output "swap-counterex.ps"
set xlabel "Utility of fixed alternative (m)"
set ylabel "Utility relative to stopping"
set xtics 0.5
set ytics 0.1
set key left top
plot [u=-2:2] [-0.3:0.3] 0 title "stop" with lines lw 2, qa(u)-q0(u) title "observe 1" with lines lw 2, qb(u)-q0(u) title "observe 2" with lines lw 2	
\end{comment}

@@

\begin{dfn}\label{dfn:context}
	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$, and a constant $\mu\in\R$,
	define $M_\mu = (S,s_0,A_s,T,R_\mu)$ to be $M$ with an additional action of known 
	value $\mu$, defined by:
	\begin{align*}
		R_\mu(s,E,s')      &= R(s,E,s') \\
		R_\mu(s,\bot,\bot) &= \max\{\mu, R(s,\bot,\bot)\}
	\end{align*}
\end{dfn}

\begin{thm}
	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$, 
	there exists a real interval $I(s)$ for every state $s\in I$ such that
	it is optimal to stop in state $s$ in $M_\mu$ iff $\mu\notin I(s)$.
	Furthermore, $I(s)$ contains $\max_i\mu_i(s)$ whenever it is nonempty.
\end{thm}

\begin{comment}
	\begin{proof}
	With $m(s) = \max_i\mu_i(s)$, the utility of a policy $\pi$ starting in state $s$ of $M_\mu$ is
	\[
		V^\pi_{M_\mu}(s) = \IE_{\pi}[-c\,N + \max(\mu,m(S_N)) \given S_0=s]
	\]
	and the utility of stopping in this state $\max(\mu,m(s_0))$.
	We wish to show that the set of $\mu$ such that
	\[
		\max_\pi \IE_{\pi}[-c\,N + \max(\mu,m(S_N)) - \max(\mu,m(S_0)) \given S_0=s] \le 0
	\]
	forms an interval.

@@

	Applying this with $v_1 = m(s_0)$ and $v_2=\IE_{\pi}[c\,N]$, and observing that the union of intervals containing
	a point is an interval containing that point, gives the result.
	\end{proof}	
\end{comment}

>>>>

\begin{example}[Non-indexability]
Consider a metalevel probability model with three actions. 
	$U_1$ is equally likely to be $-1.5$ or $1.5$ (low mean, high variance),
	$U_2$ is equally likely to be $0.25$ or $1.75$ (high mean, low variance),
	and $U_3=\lambda$ has a known value (the context).
The two computations are to observe exactly $U_1$ and $U_2$, respectively, each with cost $0.2$.
The corresponding metalevel MDP has 9 states and can be solved exactly.
\figref{fig:swap-counterex} plots $Q^*_\lambda(s_0,U_i) - Q^*_\lambda(s_0,\bot)$ as a function of the known value $\lambda$.
As the context $\lambda$ varies the optimal action inverts from observing 1 to observing 2.
Inversions like this are impossible for index policies.
\end{example}

\begin{figure}[htb]
\centering
\includegraphics[scale=0.7, trim=90 70 400 300]{swap-counterex.pdf}
\caption{Optimal Q-values of computing relative to stopping as a function of the utility of the fixed alternative.  Note the inversion
where for low $\lambda$ observing action~$1$ is strictly optimal, while for medium $\lambda$ observing action~$2$ is strictly optimal.}
\label{fig:swap-counterex}
\end{figure}

@@

	set terminal postscript enhanced linewidth 2
	set size 0.5,0.5
	set output "swap-counterex.ps"
	set xlabel "Utility of fixed alternative ({/Symbol l})"
	set ylabel "Utility relative to stopping"
	set xtics 0.5
	set ytics 0.1
	set key left top
	plot [u=-2:2] [-0.3:0.3] 0 title "stop" with lines lw 2, qa(u)-q0(u) title "observe 1" with lines lw 2, qb(u)-q0(u) title "observe 2" with lines lw 2	
\end{comment}

@@

\begin{dfn}\label{dfn:context}
	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$, and a constant $\lambda\in\R$,
	define $M_\lambda = (S,s_0,A_s,T,R_\lambda)$ to be $M$ with an additional action of known 
	value $\lambda$, defined by:
	\begin{align*}
		R_\lambda(s,E,s')      &= R(s,E,s') \\
		R_\lambda(s,\bot,\bot) &= \max\{\lambda, R(s,\bot,\bot)\}
	\end{align*}
\end{dfn}

Note this is equivalent to adding an extra arm with constant value $U_{k+1}=\lambda$.

\begin{thm}
	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$, 
	there exists a real interval $I(s)$ for every state $s\in S$ such that
	it is optimal to stop in state $s$ in $M_\mu$ iff $\mu\notin I(s)$.
	Furthermore, $I(s)$ contains $\max_i\mu_i(s)$ whenever it is nonempty.
\end{thm}

\begin{hiddenproof}
	\begin{proof}
	With $m(s) = \max_i\mu_i(s)$, the utility of a policy $\pi$ starting in state $s$ of $M_\mu$ is
	\[
		V^\pi_{M_\mu}(s) = \IE_{\pi}[-c\,N + \max(\mu,m(S_N)) \given S_0=s]
	\]
	and the utility of stopping in this state $\max(\mu,m(s_0))$.
	We wish to show that the set of $\mu$ such that
	\[
		\max_\pi \IE_{\pi}[-c\,N + \max(\mu,m(S_N)) - \max(\mu,m(S_0)) \given S_0=s] \le 0
	\]
	forms an interval.

@@

	Applying this with $v_1 = m(s_0)$ and $v_2=\IE_{\pi}[c\,N]$, and observing that the union of intervals containing
	a point is an interval containing that point, gives the result.
	\end{proof}	
\end{hiddenproof}

### 2012-04-03 paper/vago.uai/introduction.tex
as submitted

The broad family of sequential decision problems includes
combinatorial search problems, game playing, robotic path planning,
model-predictive control problems, Markov decision processes (MDP), whether fully or
partially observable, and a huge range of applications. In almost all
realistic instances, exact solution is intractable and approximate
methods are sought. Perhaps the most popular approach is to simulate
a limited number of possible future action sequences, 
in order to find a move in the current state that is (hopefully)
near-optimal. One of the most successful
examples of this approach are sampling algorithms for
playing the game of Go \citep{Gelly.mogo} based on the upper-confidence bound (UCB)
method \citep{Kocsis+Szepesvari:2006}. This paper examines the issue of sampling for move selection from
first principles, by defining an appropriate meta-level MDP and formally
analyzing some of its properties. Although
the meta-level MDP is also intractable, approximation schemes based on
provable value of information bounds for this problem are presented. 
These lead to practical sampling schemes that
are promising: they out-perform the UCB-based schemes,
as indicated by empirical results on both ``flat'' selection and
game-playing in Go.

As selection is extremely general \citep{TolpinShimony:2012}, we focus on the adversarial
game-playing domain as an example for clarity.
A typical game-playing algorithm explores a tree or graph of action sequences
with some limit on depth, using pruning methods to avoid irrelevant
subtrees; based on what it finds in the explored portion of the state space, the algorithm
then selects a move. This exploration is repeated after observing the move
made by the opponent (whether or not information from previous explorations
is reused is discussed in Section \ref{sec:control-redistribution}).

Clearly, it is desirable to select the best possible move using the
least possible amount of exploration. For a given amount of
exploration, decision quality can be improved by directing exploration
towards those actions sequences whose outcomes are helpful in selecting
a good move. Thus, the {\em metalevel} decision problem is to choose
what future action sequences to explore (or, more generally, what
deliberative computations to do), while the {\em object-level}
decision problem is to choose an action to execute in the real world.

That the metalevel decision problem can itself be formulated and
solved decision-theoretically was noted by \citet{Matheson:1968},
borrowing directly from the related concept of {\em information value
  theory}~\citep{Howard:1966}. In essence, computations can be
selected according to the expected improvement in decision quality
resulting from their execution. I. J.~\citet{Good:1968} independently
proposed using this idea to control search in chess, and later defined
``Type II rationality'' to refer to agents that optimally solve the
metalevel decision problem before acting. As interest in probabilistic
and decision-theoretic approaches in AI grew during the 1980s, several
authors explored these ideas further~\citep{Dean+Boddy:1988,Doyle:1988,Fehling+Breese:1988,Horvitz:1987b}.
Work by
\citet{Russell+Wefald:1988a,Russell+Wefald:1991a,Russell+Wefald:1991b}
formulated the metalevel sequential decision problem, employing an
explicit model of the results of computational actions, and applied
this to the control of game-playing search in Othello with encouraging
results.

An independent thread of research on metalevel control began with work
by \citet{Kocsis+Szepesvari:2006} on the UCT algorithm, which operates
in the context of {\em Monte Carlo tree search} (MCTS) algorithms.
In MCTS, each computation takes the form
of a simulating a randomized sequence of actions leading from a leaf of the
current tree to a terminal state. UCT is primarily a method for
selecting a leaf from which to conduct the next simulation, and
forms the core of the successful \textsc{MoGo} algorithm for Go 
playing \citep{Gelly+Silver:2011}.  The UCT algorithm is
based on the the theory of bandit problems \citep{Berry+Fristedt:1985} and the near-optimal
UCB1 bandit algorithm \citep{Auer+et+al:2002}. UCT applies
UCB1 recursively to select actions to perform within simulations.

It is natural to consider whether the two independent threads are
consistent; for example, are bandit algorithms such as UCB1 approximate
solutions to some particular case of the metalevel decision problem
defined by Russell and Wefald? The answer, perhaps surprisingly, is no.
The essential difference is that, in bandit problems, every trial involves
executing a real object-level action with real costs, whereas in the metareasoning
problem the trials are {\em simulations} whose cost is usually independent of the
utility of the action being simulated. 
Hence, in UCT, bandit algorithms are applied to problems that are not bandit problems.

@@

Analyzing the metalevel problem within an appropriate theoretical
framework ought to lead to more effective algorithms than those
obtained within the bandit framework.  For Monte Carlo computations,
in which samples are gathered to estimate the utilities of actions,
the metalevel decision problem is an instance of the {\em selection
problem} studied in statistics~\citep{Bechhofer:1954,Swisher+et+al:2003}.  Despite
recent work by \citet{Frazier+Powell:2010} and \citet{TolpinShimony:2012}, the theory of selection
problems is less well understood than that of bandit problems.
Accordingly, we present in \secrefs{sec:optimal}{sec:context}
a number of results concerning optimal policies for the general case
and for specific sampling distributions. In particular, we give finite
bounds on the number of samples collected by optimal policies for
Bernoulli and Gaussian distributions. We also provide a simple
counterexample to the intuitive conjecture that an optimal policy
should not spend more on deciding than the decision is worth; in fact,
it is possible for an optimal policy to compute forever. We also show
by a simple counterexample that optimal {\em index
policies}~\citep{Gittins:1989} do not necessarily exist for selection
problems.

@@

>>>>

The broad family of sequential decision problems includes
combinatorial search problems, game playing, robotic path planning,
model-predictive control problems, Markov decision processes (MDP), whether fully or
partially observable, and a huge range of applications. In almost all
realistic instances, exact solution is intractable and approximate
methods are sought. Perhaps the most popular approach is to simulate
a limited number of possible future action sequences, 
in order to find a move in the current state that is (hopefully)
near-optimal. In this paper, we develop a theoretical framework to examine the problem of selecting
{\em which} future sequences to simulate.
We derive a number of new results concerning optimal policies for this selection problem
as well as new heuristic policies for controlling Monte Carlo simulations.
As described below, these policies outperform previously published methods for
``flat'' selection and game-playing in Go.

The basic ideas behind our approach are best explained in a familiar
context such as game playing.  A typical game-playing algorithm
chooses a move by first exploring a tree or graph of move sequences
and then selecting the most promising move based on this exploration.
Classical algorithms typically explore in a fixed order, imposing a
limit on exploration depth and using pruning methods to avoid
irrelevant subtrees; they may also reuse some previous computations
(see Section \ref{sec:control-redistribution}).  Exploring unpromising
or highly predictable paths to great depth is often wasteful; for a
given amount of exploration, decision quality can be improved by
directing exploration towards those actions sequences whose outcomes
are helpful in selecting a good move. Thus, the {\em metalevel}
decision problem is to choose what future action sequences to explore
(or, more generally, what deliberative computations to do), while the
{\em object-level} decision problem is to choose an action to execute
in the real world.

That the metalevel decision problem can itself be formulated and
solved decision-theoretically was noted by \citet{Matheson:1968},
borrowing from the related concept of {\em information value
  theory}~\citep{Howard:1966}. In essence, computations can be
selected according to the expected improvement in decision quality
resulting from their execution. I. J.~\citet{Good:1968} independently
proposed using this idea to control search in chess, and later defined
``Type II rationality'' to refer to agents that optimally solve the
metalevel decision problem before acting. As interest in probabilistic
and decision-theoretic approaches in AI grew during the 1980s, several
authors explored these ideas further~\citep{Dean+Boddy:1988,Doyle:1988,Fehling+Breese:1988,Horvitz:1987b}.
Work by
\citet{Russell+Wefald:1988a,Russell+Wefald:1991a,Russell+Wefald:1991b}
formulated the metalevel sequential decision problem, employing an
explicit model of the results of computational actions, and applied
this to the control of game-playing search in Othello with encouraging
results.

An independent thread of research on metalevel control began with work
by \citet{Kocsis+Szepesvari:2006} on the UCT algorithm, which operates
in the context of {\em Monte Carlo tree search} (MCTS) algorithms.
In MCTS, each computation takes the form
of a simulation of a randomized sequence of actions leading from a leaf of the
current tree to a terminal state. UCT is primarily a method for
selecting a leaf from which to conduct the next simulation, and
forms the core of the successful \textsc{MoGo} algorithm for Go 
playing \citep{Gelly+Silver:2011}.  The UCT algorithm is
based on the the theory of bandit problems \citep{Berry+Fristedt:1985} and the asymptotically near-optimal
UCB1 bandit algorithm \citep{Auer+et+al:2002}. UCT applies
UCB1 recursively to select actions to perform within simulations.

It is natural to consider whether the two independent threads are
consistent; for example, are bandit algorithms such as UCB1
approximate solutions to some particular case of the metalevel
decision problem defined by Russell and Wefald? The answer, perhaps
surprisingly, is no.  The essential difference is that, in bandit
problems, every trial involves executing a real object-level action
with real costs, whereas in the metareasoning problem the trials are
{\em simulations} whose cost is usually independent of the utility of
the action being simulated.  Hence, as \citet{Audibert+al:2010}
and \citet{Bubeck+al:2011} have also noted, UCT applies bandit
algorithms to problems that are not bandit problems.

@@

Analyzing the metalevel problem within an appropriate theoretical
framework ought to lead to more effective algorithms than those
obtained within the bandit framework.  For Monte Carlo computations,
in which samples are gathered to estimate the utilities of actions,
the metalevel decision problem is an instance of the {\em selection
problem} studied in statistics~\citep{Bechhofer:1954,Swisher+et+al:2003}.  Despite
some recent work \citep{Frazier+Powell:2010,TolpinShimony:2012}, the theory of selection
problems is less well understood than that of bandit problems.
Most work has focused on the probability of selection error rather than
optimal policies~\citep{Bubeck+al:2011}.
Accordingly, we present in \secrefs{sec:optimal}{sec:context}
a number of results concerning optimal policies for the general case
as well as specific finite
bounds on the number of samples collected by optimal policies for
Bernoulli arms with beta priors. We also provide a simple
counterexample to the intuitive conjecture that an optimal policy
should not spend more on deciding than the decision is worth; in fact,
it is possible for an optimal policy to compute forever. We also show
by counterexample that optimal {\em index
policies}~\citep{Gittins:1989} may not exist for selection
problems.

@@

(Proofs omitted due to space constraints can be found in the supplementary material.)
